---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r warning=FALSE}
library(readr)
## auto unzip and read csv
activity <- read_csv("repdata_data_activity.zip")
```


## What is mean total number of steps taken per day?
```{r}
stepsPerDay <- aggregate(steps~date, 
                         data = activity, 
                         FUN = sum, 
                         na.rm = TRUE
                         )
hist(stepsPerDay$steps,breaks = 10, main = "Histogram of total steps per day", xlab = "steps per day")
mn <- mean(stepsPerDay$steps)
md <- median(stepsPerDay$steps)
mn
md
ans <- paste("The mean total number of steps is",mn,"steps.\nThe median total number of steps is",md,"steps.")
print(ans)
```
The mean total number of steps is $`r mn`$ steps.
The median total number of steps is $`r md`$ steps.

## What is the average daily activity pattern?
see plot...
```{r}
meanStepsPerIntv <- aggregate(steps~interval, 
                              data = activity, 
                              FUN = mean, 
                              na.rm =TRUE
                              ) 
plot(steps~interval, data = meanStepsPerIntv, type = 'l', main = "average step across interval")
## time of day people walk a lot?
walkalot <- meanStepsPerIntv$interval[which.max(meanStepsPerIntv$steps)]
hh <- round(walkalot/60,0)
walkalot
ans <- paste("people walk the most steps at interval",walkalot,". This is about",hh,": 00 Time.")
print(ans)
```
people walk a lot at Interval `r walkalot`. This is roughly `r hh`:00 Time.

## Imputing missing values
### impute strategy:
First let's decide how to impute the data.

1. find which index need to be imputed
2. find interval of that index
3. get the mean of step on that interval across every day
4. impute the steps in that index
```{r}
## will impute with mean of each interval, mean taken across all days
NAs <- !complete.cases(activity) 
NAindices <- which(NAs)

imputedActivity <- activity
for (i in NAindices){
    missingInterval <- activity$interval[i]
    imputedActivity$steps[i] <- meanStepsPerIntv$steps[meanStepsPerIntv$interval == missingInterval]
}
## count errors before and after
errorCount1 <- sum(NAs)
errorCount2 <- sum(!complete.cases(imputedActivity))
ans <- paste("There was",errorCount1,"errors before imputing and",errorCount2,"error after imputing.")
print(ans)
```
there was `r errorCount1` errors before imputing. After impute there is `r errorCount2` error.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
isweekday <- function(dte){
    ifelse(
        weekdays(dte) %in% c("Sunday", "Saturday"),
        FALSE,
        TRUE)
    }
daytype <- as.factor(
    ifelse(
        isweekday(imputedActivity$date),
        "weekday",
        "weekend"
        )
    )
imputedActivity <- cbind(imputedActivity, daytype)

agg<-aggregate(steps~interval+daytype,data = imputedActivity, FUN = sum )
par(mfrow=c(2,1),mar=c(4,2,1,1))
plot(steps~interval, 
            data = subset(agg,daytype == "weekday"), 
            type='l',
            main = "weekday"
            )
plot(steps~interval,
            data = subset(agg,daytype == "weekend"), 
            type='l',
            main = "weekend")

```
There are differences in number of steps and whether it is weekday or weekend. people walk more on weekday than on weekend (notice the different y-scale). Also, number of steps shoot up on earlier interval on weekday than on weekend. This tells how people getting to work on weekday and rest on weekend affect our data.  
